Measurement Bias from Address Aliasing

### 2014-05-18 paper.tex
New mitigate alias section at the end
Avoid focus on compilers, not the point of the paper.

% Table with detailed performance data, use estimated for O2 because it does not need as many columns (!)
\begin{table*}[t]
  \caption{Most interesting performance counters and correlation with cycle count for optimization O2, estimated cost accounting for constant overhead.}
  \label{tab:convstats}

  %\renewcommand{\tabcolsep}{3pt} % Less padding between cells
  \pgfplotstabletypeset[
    font=\small,
    int detect, % Output whole numbers for counter values
    col sep=comma,
    columns={Performance counter, Correlation, 0, 2, 4, 8},
    column type=r,
    columns/Performance counter/.style={
      string type, 
      column type=l,
      postproc cell content/.append code={
        \pgfkeysalso{@cell content=\perfctr{##1}}
      }
    },
    columns/Correlation/.style={
      fixed,
      fixed zerofill,
      precision=2
    },
    every head row/.style={
      output empty row,
      before row={\toprule
        Performance counter & $r$ & 0 & 2 & 4 & 8 \\
        },
      after row=\midrule,
    },
    every last row/.style={after row=\bottomrule}
  ]{bin/conv-default-o2.estimate.csv}
\end{table*}

@@

\subsection{Architectural Optimization}
It is perhaps unfair to assume the compiler can fix this when only given a generic optimization flag, as aliasing effects is (presumably) an Intel-specific issue.
Another option is to utilize the C99 keyword \emph{restrict} to explicitly tell the compiler that the buffers are disjoint in memory.

\begin{lstlisting}[language=C]
void conv( int n,
            const float * restrict in,
            const float * restrict out );
\end{lstlisting}

Even with the updated signature compiled with -O3 -march=native, there is still room for improvement by inserting manual offsets.

@@

Most malloc alternatives focuses heavily on performance in multithreaded environments.
Heap allocation is intrinsically inefficient in that all threads share the same address space, leading to a high potential lock contention on memory accesses.
To our knowledge, there are no commonly used allocators that specifically tries to mitigate aliasing in heap allocated memory.
This despite explicit recommendation from Intel that special purpose allocators should care about aliasing for optimization. [cite]

@@

@@

%\onecolumn

>>>>

% Table with detailed performance data, use estimated for O2 because it does not need as many columns (!)
\begin{table*}[t]
  \caption{Most interesting performance counters and correlation with cycle count for optimization O2, estimated cost accounting for constant overhead.}
  \label{tab:convstats}
  \pgfplotstabletypeset[
    font=\small,
    int detect, % Output whole numbers for counter values
    col sep=comma,
    columns={Performance counter, Correlation, 0, 2, 4, 8},
    column type=r,
    columns/Performance counter/.style={
      string type, 
      column type=l,
      postproc cell content/.append code={
        \pgfkeysalso{@cell content=\perfctr{##1}}
      }
    },
    columns/Correlation/.style={
      fixed,
      fixed zerofill,
      precision=2
    },
    every head row/.style={
      output empty row,
      before row={\toprule
        Performance counter & $r$ & 0 & 2 & 4 & 8 \\
        },
      after row=\midrule,
    },
    every last row/.style={after row=\bottomrule}
  ]{bin/conv-default-o2.estimate.csv}
\end{table*}

@@

@@

Most malloc alternatives focuses heavily on performance in multithreaded environments.
Heap allocation is intrinsically inefficient in that all threads share the same address space, leading to a high potential lock contention on memory accesses.

@@

\subsection{Ways to Deal with Address Aliasing}
Measurement bias from address aliasing can be significant in real applications, with potentially large performance penalties.
However, with an understanding of the underlying mechanism that causes bias, the effects can to some extent be predicted and accounted for in software.
The most relevant scenario to consider is probably code that can hit the aliasing \texttt{mmap} scenario, where performance impact can be consistently bad over all environments.
In our convolution kernel example, there are several ways to limit the extent of aliasing.

\paragraph{Mark buffers with \texttt{restrict}}
In our convolution kernel implementation, the compiler has to account for the fact that input and output pointers might alias, or that the buffers partially overlap.
This limits the extent generated code can keep data in registers without updating the values from memory, as a write to one buffer potentially could invalidate a cached value from the other.
The C99 keyword \emph{restrict} can be used to explicitly tell the compiler that accesses through a pointer does not alias with any other, allowing for more efficient code generation with fewer memory accesses.
\begin{lstlisting}[language=C]
void conv( int n,
            const float * restrict input,
            float * restrict output );
\end{lstlisting}
With this updated function prototype, most of the address aliasing events are gone for the default alignment.

\paragraph{Use a special purpose allocator}
We found that heap allocators are prone to generate pairwise aliasing buffers, the worst case for functions such as the convolution example.
A potential solution could be to apply some heuristic to randomize addresses more, and in particular not always return the same 12 bit suffix for large allocations.
This goes somewhat agains conventional wisdom that more alignment is better.
The Intel optimization manual actually mentiones this in \emph{User/Source Coding Rule 8}, suggesting that special purpose allocators could be used to avoid aliasing ~\cite{OptimizationManual}.
However, to our knowledge there are no commonly used allocators that specifically tries to mitigate aliasing in heap allocated memory.

\paragraph{Manually adjust memory}
In some cases it might make sense to explicitly control the memory addresses used, for example forcing some fixed relative offset between input and output pointers.
This can be achieved by exploiting the fact the \texttt{mmap} is in fact guaranteed to be placed at a page boundary, and use that directly instead of \texttt{malloc}.
The following approach can be used to make an anonymous memory mapping with offset \texttt{d} bytes away from page alignment.
\begin{lstlisting}[breaklines=true]
    mmap(NULL, (n + d), PROT_READ | PROT_WRITE, MAP_PRIVATE | MAP_ANONYMOUS, -1, 0) + d;
\end{lstlisting}
The same pointer difference must of course by subtracted before unmapping the memory.

@@
